---
title: "Assignment 4"
author: "Nicky Bell"
output: 
  html_document:
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  include = FALSE
)
```

### Assignment Instructions

You must submit both a .Rmd file and knitted output in either HTML or PDF format. In addition, you **must** `echo` your code and hide any messages and warnings in your knitted output to assist the TAs in grading your work. Assignments missing either file, do not `echo` the code, or do not hide messages and warnings will receive a 2 point deduction.

Where you are asked to interpret the results of our analysis, you must include a brief written response to receive full credit for the question.

Please review the course policy on Generative AI tools in the [syllabus](https://canvas.upenn.edu/courses/1867281/assignments/syllabus).

### Part I: 1970 Vietnam War Draft Lottery

On December 1, 1969, the U.S. Selective Service held a lottery to determine the order in which young men would be drafted to serve in the Vietnam War. To conduct the lottery, each day of the year was written on a slip of paper, and the slips were placed in plastic capsules. For each month starting with January, the capsules were mixed in a shoebox and then dumped in a large glass container. Finally, the large glass container was mixed, and the capsules were drawn one at a time. The order in which the days were drawn determined the order of call to military service.

The lottery results soon proved controversial when statisticans claimed that the lottery was not truly random. As evidence, they pointed out that the later months of the year tended to have lower draft numbers than the earlier months. (The expected average draft number in any month is 183.5.)

1. Load the data (`1970_vietnam_draft_lottery.csv`). Using a loop, conduct a t-test comparing the mean draft numbers for each month to the expected mean of 183.5. Save the mean draft number and the confidence interval for each month in a new data frame. 

```{r}
library(tidyverse)
draft <- read_csv(here::here("data/week4/1970_vietnam_draft_lottery.csv"))
draft_results <- tibble()
for (i in 1:12) {
  df <- filter(draft, month == i)
  t_test_result <- t.test(df$lottery_number, mu = 183.5)
  month_summary <- tibble(
    month = i,
    mean_draft_number = t_test_result$estimate,
    ci_lower = t_test_result$conf.int[1],
    ci_upper = t_test_result$conf.int[2]
  )
  draft_results <- bind_rows(draft_results, month_summary)
}
```

2. Create a plot showing the mean draft number for each month with error bars representing the 95% confidence interval. Add a horizontal line at 183.5 to indicate the expected mean draft number.

```{r}
ggplot(data = draft_results, aes(x = factor(month), y = mean_draft_number)) +
  geom_col() +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2) +
  geom_hline(yintercept = 183.5, linetype = "dashed", color = "red") +
  scale_x_discrete(labels = month.abb) +
  labs(
    title = "Mean Draft Number by Month with 95% Confidence Intervals",
    x = "Month",
    y = "Mean Draft Number"
  ) +
  theme_minimal()
```

3. Based on your results, do you think the lottery was random? Explain your reasoning.

### Part II: Predicting the 2024 Presidential Election

The Cooperative Election Study (CES) is the largest academic survey focused on American elections. Each election year, more than 50,000 Americans are interviewed before and after the election. The full data is released to the public months after the election, but the CES team releases summary results of respondents' vote intention in key battleground elections just before the election.

For this section, you will need to install and load the `{survey}` and `{srvyr}` package.

4. Load the data (`ces24.csv`). Limit the data to respondents who already voted or plan to vote (`CC24_363` is 1, 2, 3, or 4). In addition, create a new variable called `harris` which is 1 if the respondent voted (`CC24_364a == 1`) or intended to vote (`CC24_364b == 1`) for Kamala Harris and 0 otherwise. (*Hint: Use a `case_when()` function rather than an `ifelse()` function to create this variable.*)

    Finally, create a [survey design object](https://tidy-survey-r.github.io/tidy-survey-book/c04-getting-started.html#setup-des-obj) from your modified data frame using the code below:

```{r, echo = TRUE, eval = FALSE}
ces_design <- as_survey_design(ces_mod, weights = commonweight)
```

```{r}
library(survey)
library(srvyr)
ces <- read_csv(here::here("data/week4/ces24.csv"))
ces_mod <- ces |>
  filter(CC24_363 %in% c(2, 3, 4, 5)) |>
  mutate(
    harris = case_when(CC24_364a == 1 ~ 1, CC24_364b == 1 ~ 1, TRUE ~ 0)
  )
ces_design <- as_survey_design(ces_mod, weights = commonweight)
```

5. Using a loop, conduct a t-test comparing the mean proportion of respondents who voted or intend to vote for Kamala Harris (your new `harris` variable) in each swing state to 0.5. Each state is represented by a numeric code in the variable `inputstate`:

    * Arizona (4)
    * Georgia (13)
    * Michigan (26)
    * Nevada (32)
    * North Carolina (37)
    * Pennsylvania (42)
    * Wisconsin (55)

    I've provided some code to run a *weighted* t-test on the survey design object you created earlier to ensure that our results reflect a representative sample of American voters. You will need to modify it to loop over the states and save the results in a new data frame.

    *Hint: Look at the formula in the `svyttest()` function below. We are comparing the mean of `harris - .5` to `0`; given this, think carefully about how you should interpret positive and negative means that are estimated by the weighted t-test function.*

```{r, echo = TRUE, eval = FALSE}
svyttest(harris - .5 ~ 0, design = ces_design, na.rm = TRUE)
```

```{r}
harris_results <- tibble()
for (i in c(4, 13, 26, 32, 37, 42, 55)) {
  state_design <- filter(ces_design, inputstate == i)
  t_test_result <- svyttest(
    harris - .5 ~ 0,
    design = state_design,
    na.rm = TRUE
  )
  state_summary <- tibble(
    state = i,
    mean_harris = t_test_result$estimate,
    p_value = t_test_result$p.value
  )
  harris_results <- bind_rows(harris_results, state_summary)
}
knitr::kable(
  harris_results,
  col.names = c("State", "Mean Proportion for Harris", "P-value"),
  caption = "Weighted T-Test Results for Kamala Harris by State"
)
```

6. Based on your results, how did the CES perform in predicting the 2024 presidential election? Explain your reasoning.